A universal of human social cognition: Children from 17 communities process gaze in similar ways

### 2024-10-18 supplement/supplement.Rmd
paper long version review 1

```{r demtable, echo=F}
demtab <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent)

kableExtra::kable(demtab,
             col.names = c("Community", "Country", "Continent", "N", "Sex (male)", "Age (range)"),
              caption = "Participant demographics.", 
             format="latex", booktabs=TRUE
             )
```

A total of `r sum(as.numeric(demtab$N))` children between `r round(min(data$ageinyears), 2)` and `r round(max(data$ageinyears), 2)` provided data for the study. Children lived in `r n_distinct(data$community)` different communities, located in `r n_distinct(demtab$country)` different countries. Table \@ref(tab:demtable) gives the sample size per community together with some basic demographic information. The recruitment strategy for each community is reported in the respective site descriptions. Figure \@ref(fig:figpred1)A shows the age distribution in each setting. For some children, the exact birthday was unknown. In such cases, we set the birthday to the 30th of June of the year that would make them fall into the reported age category.

## Site descriptions

@@

@@

We adapted the task developed by Prein and colleagues [@prein2023tango]. We refer to the original publication for a detailed description of its development, implementation and psychometric evaluation (in Germany). Below, we give an overview and focus on the cross-cultural adaptation of the task. The task itself, including all the versions used in the study, can be accessed via the following website: https://ccp-odc.eva.mpg.de/tango-cc/

@@

# Analysis

@@

We pre-registered an analysis plan at <!--https://osf.io/tdsvc/?view_only=f404d66c967542f28ce4af16371f34e6-->https://osf.io/tdsvc. The analyses reported below follow the same logic as the pre-registered analyses but they differ substantially in detail. The main difference lies in the random effects structure we used in the models. In the pre-registration, we did not include random slopes for target centrality (i.e. the distance of the location where the balloon landed from the center of the screen), neither within subject nor within cultural setting. Instead, we included random slopes for trial within subject. We decided to change this and remove random slopes for trial but include them for target centrality. The reasons were as follows: trial effects in the sense of learning across trials were unlikely because participants did not receive differential feedback. Furthermore, in multiple studies with the same task since the registration we never found any trial effects [@prein2023tango; @prein2023variation]. We therefore removed trial from the models completely. On the other hand, we included random slopes for target centrality to be able to look at cross-cultural variation. The cognitive model was not mentioned in the pre-registration because it is a more recent development [see @prein2023variation].

@@

```{r}
# mdata <- data%>%
#   ungroup()%>%
#   mutate(ageinyears = as.numeric(scale(ageinyears)),
#          targetcentralityx = as.numeric(scale(targetcentralityx)),
#          click = abs(clickdistfromtargetcenterx))

@@

The second analysis focused on signatures in the data that suggest that children in all cultural settings use similar cognitive processes to infer the location of the balloon. The first -- somewhat weaker -- signature was simply an increase in performance with age. The second -- stronger -- signature was a strong effect of target centrality. That is, the further away from the center the balloon lands, the more imprecise should participants be. This second signature is derived from the cognitive model introduced in @prein2023variation. It predicts that trials in which the agent looks further away from the center (i.e., to the left or right side of the screen) result in higher levels of imprecision compared to trials in which the agent looks closer to the middle.

@@

@@

```{r figdev, include = T, fig.align = "center", fig.cap = "A) Developmental trajectory across and B) by community. The developmental trajectories are predicted based on a model of the data aggregated for each participant. C) Performance by screen section across and D) by community. Each bin covers 1/10th of the screen. Points show means, and error bars 95% confidence intervals for the data within that bin aggregated across participants. Transparent dots in A) and C) show aggregated data for each individual.", out.width="100%"}

# pvis_pred <- plot_grid(
# plot_grid(
#   pagg_dev, pind, rel_widths = c(1,1.5), align = "h", axis = "bt",labels = c("A", "B")
# ),
# plot_grid(
#   ptagg, ptind, rel_widths = c(1,1.5), align = "h", axis = "bt", labels = c("C", "D")
# ), nrow = 2, align = "v"
# )
# 
# ggsave(plot = pvis_pred, "../figures/pvis_pred.png", width = 12, height = 8, bg = "white", scale = 1.25)

@@

The vector-based gaze estimation model has been described in detail in @prein2023variation. In brief, it inversely models the process generating clicks based on observed eye movements. Formally, the model is defined as:

@@

```{r}
# scdata <- mdata%>%
#   group_by(subjid, community, ageinyears, touchscreen, household, children, younger_children)%>%
#   summarise(mean_click = mean(click))%>%
#   ungroup()%>%
#   group_by(community)%>%
#   mutate(
#          household = scale(household),
#          children = scale(children),
#          younger_children = scale(younger_children)
#          )%>%
#   na.omit()
# 
# bm_pred_null <- brm(mean_click ~ ageinyears + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc <- brm(mean_click ~ ageinyears + touchscreen + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8))%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_hh <- brm(mean_click ~ ageinyears + touchscreen + household + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# 
# bm_pred_tc_c <- brm(mean_click ~ ageinyears + touchscreen + children + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_yc <- brm(mean_click ~ ageinyears + touchscreen + younger_children + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_rand <- brm(mean_click ~ ageinyears + touchscreen + (touchscreen + ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
```

@@

The final analysis focused on whether we could predict performance in the task by aspects of everyday experience. For the ease of model fitting, we aggregated the data for each participant so that models predicted the average imprecision across trials [@prein2023variation showed that the performance mean is nearly perfectly correlated with $\sigma_v$. When re-running the winning model on the trial-by-trial data, the results were qualitatively and quantitatively identical.]. The aspects we could focus on were limited to the questions we asked in the questionnaire. That is, we analysed the effects of exposure to technology and household composition.

@@

```{r predcomptable, echo=F}
pred_mocel_comp <- readRDS("../saves/model_comp.rds")%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  select(model, elpd_diff, se_diff, waic, se_waic, value)%>%
  mutate(model = recode(model, 
                        bm_pred_null = "null",
                        bm_pred_tc = "touchscreen",
                        bm_pred_tc_hh = "household",
                        bm_pred_tc_c = "children",
                        bm_pred_tc_yc = "younger children",
                        bm_pred_tc_rand = "touchscreen (by culture)"))

kableExtra::kable(pred_mocel_comp,
             col.names = c("Model","diff\\textsubscript{WAIC}", "diff\\textsubscript{SE}", "WAIC", "SE", "Weight"),
             caption = "Log Bayes Factors in favor of the gaze model compared to two alternative models.",
             format="latex", booktabs=TRUE, escape = F
             )

Table \@ref(tab:predcomptable) shows that the model predicting average imprecision by the availability of touchscreens performed best and better than any of the models including a predictor referring to household composition. However, the difference between the models was very small (`ELPD` < 1, see Table \@ref(tab:predcomptable)), suggesting that they are largely equivalent, and, for the sake of consistency, we therefore inspected the posterior estimates for household composition predictors nevertheless.

@@

>>>>

# Overview

This document supplements the paper "A universal of human social cognition: Children from 17 communities process gaze in similar ways". Some text passages and figures are the same in the supplementary material and the main paper. This redundancy is intended and ensures that the supplementary material is self-contained and readable.

This document is structured as follows: we first provide a detailed description of the different [study sites](#site-descriptions), inlcuding information about geography, history, subsistence, socialization goals and recruitment strategy. Next we summarize the resulting [sample](#participants). Then we go on to provide additional details about the [methods](#method) and the [analysis and results](#analysis-and-results). The online version in the repository (`.Rmd` file) includes the analysis code for the analysis reported in the paper and the supplementary material. In the [appendix](#appendix) we provide the English version of the voice overs children heard during the study.

# Site descriptions

@@

```{r demtable, echo=F}
demtab <- bind_rows(

data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste(mean_age, " (", min_age, " - ", max_age, ")"),
         age_group = "total")%>%
  select(-c(mean_age, min_age, max_age)),

data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent, age_group)%>%
  summarise(N = n_distinct(subjid),
            m = sum(sex == "m", na.rm = T))%>%
  mutate(age_group = paste0(age_group, " years"))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)

)%>%
  arrange(continent, community)%>%
  select(continent, country, community, age_group, N, m, age)%>%
  ungroup()%>%
  mutate(continent = as.character(continent),
         country = replace(country, duplicated(country), " "),
         community = replace(community, duplicated(community), " "),
         continent = replace(continent, duplicated(continent), " "))

kableExtra::kable(demtab,
             col.names = c("Continent","Country","Community", "Age group","N", "Sex (male)", "Age (range)"),
              caption = "Participant demographics.", 
             format="latex", booktabs=TRUE, longtable = TRUE
             )%>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))
```

A total of `r sum(as.numeric(demtab$N))` children between `r round(min(data$ageinyears), 2)` and `r round(max(data$ageinyears), 2)` provided data for the study. Children lived in `r n_distinct(data$community)` different communities, located in `r n_distinct(demtab$country)` different countries. Table \@ref(tab:demtable) gives the sample size per community together with some basic demographic information. The recruitment strategy for each community is reported in the respective site description above. Figure \@ref(fig:figpred1)A shows the age distribution in each setting. For some children, the exact birthday was unknown. In such cases, we set the birthday to the 30th of June of the year that would make them fall into the reported age category.

@@

We adapted the task developed by Prein and colleagues [@prein2024tango]. We refer to the original publication for a detailed description of its development, implementation and psychometric evaluation (in Germany). Below, we give an overview and focus on the cross-cultural adaptation of the task. The task itself, including all the versions used in the study, can be accessed via the following website: https://ccp-odc.eva.mpg.de/tango-cc/

@@

# Analysis and results

@@

@@

```{r}
mdata <- data%>%
  ungroup()%>%
  mutate(ageinyears = as.numeric(scale(ageinyears)),
         targetcentralityx = as.numeric(scale(targetcentralityx)),
         click = abs(clickdistfromtargetcenterx))

@@

The second analysis focused on signatures in the data that suggest that children in all cultural settings use similar cognitive processes to infer the location of the balloon. The first -- somewhat weaker -- signature was simply an increase in performance with age. The second -- stronger -- signature was a strong effect of target centrality. That is, the further away from the center the balloon lands, the more imprecise should participants be. This second signature is derived from the cognitive model introduced in @prein2024variation. It predicts that trials in which the agent looks further away from the center (i.e., to the left or right side of the screen) result in higher levels of imprecision compared to trials in which the agent looks closer to the middle.

@@

```{r}
side_bias <- data%>%
  mutate(side = ifelse(targetcenterx < 960, "left", "right"),
         error = ifelse(clickdistfromtargetcenterx < 0, "left", "right"))%>%
  group_by(side)%>%
  summarise(left_bias = mean(error == "left"),
            right_bias = mean(error == "right"))
```

Visualization of the data also showed the predicted u-shaped pattern in all communities (see Fig. \@ref(fig:figdev)C). The visualization also showed that the u shape is not symmetric. Instead, imprecision was higher when the target landed on the left side of the screen. One potential explanation for this pattern could be the wide-spread dominance of right-handedness. Remember that participants had a general center bias, that is, they were more likely to touch the screen to the side of the object that was closer to the center. Now, when touching a location on the left side of the screen, right-handed participants need to cross their hand in front of their body and might be inclined to touch the screen slightly further to the right compared to when the target lands on the right side of the screen. This might amplify the center bias. In the data, we saw that when the target landed on the left side of the screen, `r round(side_bias%>%filter(side == "left")%>%pull(right_bias),2)*100`% of touches were to the right of the target and closer to the center of the screen. On the other hand, when the target landed on the right, `r round(side_bias%>%filter(side == "right")%>%pull(left_bias),2)*100`% of touches were left of the target and closer to the center. A center bias alone would have predicted these numbers to be the same.

@@

```{r figdev, include = T, fig.align = "center", fig.cap = "A) Developmental trajectory across and B) by community. The developmental trajectories are predicted based on a model of the data aggregated for each participant. C) Performance by screen section across and D) by community. Each bin covers 1/10th of the screen. Points show means, and error bars 95\\% confidence intervals for the data within that bin aggregated across participants. Transparent dots in A) and C) show aggregated data for each individual.", out.width="100%"}

# pvis_pred <- plot_grid(
# plot_grid(
#   pagg_dev, pind, rel_widths = c(1,1.75), align = "h", axis = "bt",labels = c("A", "B")
# ),
# plot_grid(
#   ptagg, ptind, rel_widths = c(1,1.75), align = "h", axis = "bt", labels = c("C", "D")
# ), nrow = 2, align = "v"
# )
# 
# ggsave(plot = pvis_pred, "../figures/pvis_pred.png", width = 12, height = 8, bg = "white", scale = 1.25)

@@

The vector-based gaze estimation model has been described in detail in @prein2024variation. In brief, it inversely models the process generating clicks based on observed eye movements. Formally, the model is defined as:

@@

```{r}
# scdata <- mdata%>%
#   group_by(subjid, community, ageinyears, touchscreen, household, children, younger_children)%>%
#   summarise(mean_click = mean(click))%>%
#   ungroup()%>%
#   group_by(community)%>%
#   mutate(
#          household = scale(household),
#          children = scale(children),
#          younger_children = scale(younger_children)
#          )%>%
#   na.omit()
# 
# bm_pred_null <- brm(mean_click ~ ageinyears + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              #backend = "cmdstanr",
#              #threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc <- brm(mean_click ~ ageinyears + touchscreen + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              #backend = "cmdstanr",
#              #threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_hh <- brm(mean_click ~ ageinyears + touchscreen + household + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95)#,
#              #backend = "cmdstanr",
#              #threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# 
# bm_pred_tc_c <- brm(mean_click ~ ageinyears + touchscreen + children + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95)#,
#              #backend = "cmdstanr",
#              #threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_yc <- brm(mean_click ~ ageinyears + touchscreen + younger_children + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95)#,
#              #backend = "cmdstanr",
#              #threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))

# bm_pred_tc_rand <- brm(mean_click ~ ageinyears + touchscreen + (touchscreen + ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
```

@@

The final analysis focused on whether we could predict performance in the task by aspects of everyday experience. For the ease of model fitting, we aggregated the data for each participant so that models predicted the average imprecision across trials [@prein2024variation showed that the performance mean is nearly perfectly correlated with $\sigma_v$. When re-running the winning model on the trial-by-trial data, the results were qualitatively and quantitatively identical.]. The aspects we could focus on were limited to the questions we asked in the questionnaire. That is, we analysed the effects of exposure to technology and household composition.

@@

```{r predcomptable, echo=F}
pred_model_comp <- readRDS("../saves/model_comp.rds")%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  select(model, elpd_diff, se_diff, waic, se_waic, value)%>%
  mutate(model = recode(model, 
                        bm_pred_null = "null",
                        bm_pred_tc = "touchscreen",
                        bm_pred_tc_hh = "touchscreen + household",
                        bm_pred_tc_c = "touchscreen + children",
                        bm_pred_tc_yc = "touchscreen + younger children",
                        bm_pred_tc_rand = "touchscreen (by culture)"))

kableExtra::kable(pred_model_comp,
             col.names = linebreak(c("Model","diff\\textsubscript{WAIC}", "diff\\textsubscript{SE}", "WAIC", "SE\\textsubscript{WAIC}", "Weight")),
             caption = "Comparison of models predicting individual-level variation.",
             format="latex", booktabs=TRUE, escape = F
             )
```

Table \@ref(tab:predcomptable) shows that the model predicting average imprecision by the availability of touchscreens performed best and better than any of the models including a predictor referring to household composition. However, the difference between the models was very small ($diff_{ELPD}$ < 1, see Table \@ref(tab:predcomptable)), suggesting that they are largely equivalent, and, for the sake of consistency, we therefore inspected the posterior estimates for household composition predictors nevertheless.

@@

```{r}
# pred_est_sc <- 
#   bind_rows(
#   fixef(bm_pred_tc_hh)%>%
#   as_tibble(rownames = "term")%>%
#   filter(term == "household")%>%
#   mutate_if(is.numeric,round, 2),
#   
#     fixef(bm_pred_tc_c)%>%
#   as_tibble(rownames = "term")%>%
#   filter(term == "children")%>%
#   mutate_if(is.numeric,round, 2),
#   
#     fixef(bm_pred_tc_yc)%>%
#   as_tibble(rownames = "term")%>%
#   filter(term == "younger_children")%>%
#   mutate_if(is.numeric,round, 2)
#   
#   )
# 
# saveRDS(pred_est_sc, "../saves/pred_est_sc.rds")
# 
# pred_est_sc <- readRDS("../saves/pred_est_sc.rds")
```
